---
title: 'Basic R Notebook: Lecture 1'
author: 'Bikram'
output:
  html_document: default
  pdf_document: default
  html_notebook: default
---

Preliminaries  
Use `ctrl+alt+i` to create the R environment (thanks, Soo Han Soon)


Returns the current working directory
```{r}
getwd()
```

Provides help on a specific function
```{r}
help(getwd)
?getwd
```

Sets the current working directory
```{r}
setwd("/Users/Bikram/Google Drive/AnalyticsEdge/Week 1/R code")
```


Lists the files in a directory

```{r}
dir("/Users/Bikram/Google Drive/AnalyticsEdge/Datasets/")
```

Lists objects stored in the R workspace
```{r}
ls()
```


Assigns a number to a variable

```{r}
x<-40
x
```

(Alternative to) assigning a number to a variable. Mostly '<-' is preferred for assignment.
```{r}
x=40
x
```

Common functions - exponential, inverse, power, addition
```{r}
exp(x)
1/x
x^3
y<-x+6
y
```

Remove a variable from the workspace

```{r}
ls()
rm(y)
ls()
```
Numbers and vectors

Concatenates (combines) numbers to form a vector

```{r}
x<-c(1,-2,3,5,pi)
x
```
Accessing specific elements of the vector

```{r}
x[3]
```
Applying operations to the vector - term by term inverse, concatenate vectors, exponentiation

```{r}
1/x
exp(x)
y<-c(x,0,x)
y
```
You can overload the sum operator by recycling the shorter vector - mathematically adding vectors of different sizes are not permitted 

```{r}
x
y
x+y
```
Finding the maximum and minimum elements and identifying the location (index) of the first max and all max

```{r}
max(x)
min(y)
which.max(x)
y
which(y==min(y))
```
Other operations - sum, product, mean, variance, standard deviation
```{r}
sum(x)
prod(x)
mean(x)
sd(x)
summary(x)
```

For a vector, it provides a six number summary including min, max, mean, 1st quartile, median and 3rd quartile. This can be used with other objects too.


Maximum entry in a concatenated vector

```{r}
max(x,y)
```
Parallel maximum returns a vector of length equal to the longest argument that contains in each element, the largest element in that position of any of the vector
```{r}
?pmax
x
y
pmax(x,y)
```

Remove all variables from the workspace
```{r}
ls()
rm(list=ls())
ls()

```


Differences in the assignment using <- and ==. Try exp(a=1:5) and exp(a<-1:5)

```{r}
exp(a<-1:5)
#exp(a=1:5)
```
Generating vectors using a variety of commands

```{r}
x<--3:8
x
seq(-3,8,0.2)
rep(x,times=3)
rep(x,each=3)
```

Returns logical vector based on the check
```{r}
x
x>1

```

Dealing with missing entries 
```{r}
is.na(x)
?is.na
y<-c(x,NA)
y
is.na(y)

```

END OF LECTURE 1.