---
title: 'Basic R Notebook: Lecture 1'
author: 'Ying Xuan'
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Preliminaries

Returns the current working directory
```{r}
getwd()
```

Provides help on a specific function
```{r}
help(getwd)
# OR
?getwd
```

Sets the current working directory
```{r}
setwd("/Users/yingxuan/Documents/SUTD/Term 6/TAE/W1")
```

Lists the files in a directory
```{r}
dir("/Users/yingxuan/Documents/SUTD/Term 6/TAE/W1")
```

Lists objects stored in the R workspace
```{r}
ls()
```

Assigns a number to a variable
```{r}
x <- 1
x
```

(Alternative to) assigning a number to a variable. Mostly '<-' is preferred for assignment.
```{r}
x = 1
x
```

Common functions - exponential, inverse, power, addition
```{r}
x <- 1
exp(x)
1/x
x^3
y <- x + 6
y
```

Remove a variable from the workspace
```{r}
rm(y)
```

Numbers and vectors


Concatenates (combines) numbers to form a vector
```{r}
x <- c(1, 2, 3)
x
```

Accessing specific elements of the vector
```{r}
x <- c(1, 2, 3)
x[3]
```

Applying operations to the vector - term by term inverse, concatenate vectors, exponentiation
```{r}
x <- c(1, 2, 3)
y <- c(x, 1, x)
y 
```

You can overload the sum operator by recycling the shorter vector

* mathematically adding vectors of different sizes are not permitted 

* if one variable is only of one value, it is used as an element wise addition to all elements of the vector variable

* if both variables are vectors of different length, element wise addition still happens, and the shorter vector is recycled for element wise addition

* if both variables are vectors of the same length, addition is done according to vector element sequence

```{r}
x <- c(1, 2, 3)
y <- c(x, 1, x)
z <- x + y
z
```


Finding the maximum and minimum elements and identifying the location (index) of the first max and all max
```{r}
x <- c(1, 2, 3)
max(x) # returns value
which.max(x) # returns index
# OR
which(x == max(x)) # returns index

min(x)
which.min(x)
```


Other operations - sum, product, mean, variance, standard deviation
```{r}
x <- c(1, 2, 3)
sum(x)
prod(x)
mean(x)
sd(x)
```


For a vector, it provides a six number summary including min, max, mean, 1st quartile, median and 3rd quartile. This can be used with other objects too.
```{r}
x <- c(1, 2, 3)
summary(x)
```


Maximum entry in a concatenated vector
```{r}
x <- c(1, 2, 3)
y <- c(4, 5, 7)
max(x, y)
```


Parallel maximum returns a vector of length equal to the longest argument

- In each element of this returned vector, it returns the maximum/largest element in that index no matter which individual input vector it is from

- length of vector returned = longest input vector in the argument

- arguments are to be of either the same length, or else an argument will be fractionally recycled (re-used to give a common length across all vectors)
```{r}
?pmax
x <- c(1, -2, 3, 5, 3)
y <- c(x, 1, x)
x
y
pmax(x, y)

# vectors under comparison
# lengthened x vector VS original y vector
# 1 -2 3 5 3 1 -2 3 5 3 1
# 1 -2 3 5 3 1 1 -2 3 5 3
```


Remove all variables from the workspace
```{r}
rm(list=ls())
```



Differences in the assignment using <- and ==. Try exp(a=1:5) and exp(a<-1:5)
```{r}
exp(a <- 1:5)
# exp(a = 1:5)
```


Generating vectors using a variety of commands
```{r}
x <- -3:8 # start and end values are invlusive
x
# seq(from, to, by)
# from and to values are inclusive
seq(-3, 8, 2) # last parameter = interval
rep(x, times = 3) # times parameter = number of times to repeat (one full cycle repeated)
rep(x, each = 3) # each parameter = number of times each element to appear (together in sequence)
```



Returns logical vector based on the check
```{r}
x <- -2:3
x
x > 1
```


Dealing with missing entries 
```{r}
x <- -2:3
x
is.na(x)
y <- c(x, NA)
y
is.na(y)
```


END OF LECTURE 1.